To Airbnb?: A Question of Returns

### 2018-05-22 papers/full_paper.Rmd
remove extra viz from paper

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)

@@

After removing property level observations that are missing price data, we plot the distribution of the rates and rents.  In \autoref{rrdist} we see that most long-term weekly rentals are greater than \$200 and less than \$1,000 per week while most median nightly rates are greater than \$50 and less than \$500 per night. We filter the property level data accordingly. We also filter the short-term daily data to include only those observations with nightly rates between \$50 and \$500.

..HOLD FOR DENSITY OF RATE AND RENTS PLOT..

<!-- ```{r plot_ratesrents, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3, fig.width=5.75, fig.cap='\\label{rrdist}Distribution of Rates and Rents'} -->

<!--  load(paste0(fig.path, 'rateplot.RData')) -->
<!--  load(paste0(fig.path, 'rentplot.RData')) -->

<!--  ggMultiPlots(rate.plot, ltr.plot, cols=2) -->

@@

<!-- Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs. Inner suburbs are < 10km from the CBD, Middle suburbs 10 to 20km from CBD and Outer suburbs > 20km from CBD. In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high-end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts. The specific suburb of a property is the second, or finer scale, at which the market operates.  -->

<!-- \autoref{locmap} shows the three large submarkets (Inner, Middle and Outer) and the individual suburbs boundaries. -->

<!-- ```{r plot_suburbmap, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3.5, fig.width=3.5, fig.cap='\\label{locmap}Melbourne Submarkets'} -->

<!--  load(paste0(fig.path, 'suburbMap.RData')) -->

<!--  loc.map <- loc.map + -->
<!--    theme(axis.title.y=element_blank(), -->
<!--          axis.text.y=element_blank(), -->
<!--          axis.ticks.y=element_blank(), -->
<!--          axis.title.x=element_blank(), -->
<!--          axis.text.x=element_blank(), -->
<!--          axis.ticks.x=element_blank()) -->
<!--   ggsave(loc.map, filename=file.path(fig.path, 'loc_map.png'), -->
<!--          bg='transparent', -->
<!--          width=6, height=6) -->
<!--  # loc.map -->

<!-- ![Suburb Categories](c://dropbox/research/airBnb/figures/loc_map.png) -->

<!-- It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:  -->

<!-- 1. Rural (Outer suburbs, not Beach) -->
<!-- 2. Suburban (Middle suburbs, not Beach) -->
<!-- 3. City (Inner Suburbs, not Beach, not Core) -->
<!-- 4. City-Core (Select inner suburbs with tourist activities and near CBD) -->
<!-- 5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River) -->

<!-- We select the 18 suburbs below to represent the 'core' of the city (Table 1).  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well served by public transportation and possess abundant amenities for tourists. -->

<!-- ```{r table_corelist, results="asis", comment=FALSE, message=FALSE, echo=FALSE} -->

<!--  c.subs <- data.frame(A=core.suburbs[1:6], -->
<!--                       B=core.suburbs[7:12], -->
<!--                       C=core.suburbs[13:18]) -->
<!--  names(c.subs) <- NULL -->
<!--  rownames(c.subs) <- NULL -->

<!--  print(xtable(c.subs, caption = 'List of Core Suburbs\n'), -->
<!--        include.rownames=F, -->
<!--        caption.placement="top", -->
<!--        hline.after=c(0,6)) -->

<!-- We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short- and long-term rentals, colored by submarket, are shown in \autoref{pointmaps}. -->

<!-- Before assigning submarkets to the long-term property we must remedy the fact that about 5% of the long-term rental observations have missing latitude and longitude values.  Half of these do, however, have a latitude and longitude values for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps we apply the street centroid value to the observations with missing lat/long values.  For those without either set of lat/longs we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above. -->

<!-- ```{r load_locations, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE} -->

<!--   load(paste0(fig.path, 'ltrMap.RData')) -->
<!--   load(paste0(fig.path, 'strMap.RData')) -->

<!-- ```{r plot_locations, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=5, fig.height=10, fig.cap='\\label{pointmaps}Location Maps', fig.ext='png'} -->

<!--   str.map <- str.map + -->
<!--    theme(axis.title.y=element_blank(), -->
<!--          axis.text.y=element_blank(), -->
<!--          axis.ticks.y=element_blank(), -->
<!--          axis.title.x=element_blank(), -->
<!--          axis.text.x=element_blank(), -->
<!--          axis.ticks.x=element_blank()) -->

<!--   ltr.map <- ltr.map + -->
<!--    theme(axis.title.y=element_blank(), -->
<!--          axis.text.y=element_blank(), -->
<!--          axis.ticks.y=element_blank(), -->
<!--          axis.title.x=element_blank(), -->
<!--          axis.text.x=element_blank(), -->
<!--          axis.ticks.x=element_blank()) -->

<!--     ggsave(str.map, filename=file.path(fig.path, 'str_map.png'), -->
<!--          bg='transparent', -->
<!--          width=6, height=6) -->
<!--     ggsave(ltr.map, filename=file.path(fig.path, 'ltr_map.png'), -->
<!--          bg='transparent', -->
<!--          width=6, height=6) -->

<!--   # ggMultiPlots(str.map, ltr.map, cols=1) -->

<!-- ![Short Term Rental Locations](c://dropbox/research/airBnb/figures/str_map.png) -->

<!-- ![Long Term Rental Locations](c://dropbox/research/airBnb/figures/ltr_map.png) -->

<!--  To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (\autoref{smcounts}). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size. -->

<!-- ```{r plot_locbars, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=8.5, fig.height=4, fig.cap='\\label{smcounts}Submarket Observation Counts'} -->

<!--   load(paste0(fig.path, 'ltrsubplot.RData')) -->
<!--   load(paste0(fig.path, 'strsubplot.RData')) -->

<!--   ltrtype.map <- ltrtype.map + ylab("") + ggtitle('Long-Term Properties') -->
<!--   strtype.map <- strtype.map + ggtitle('Short-Term Properties') -->

<!--   ggMultiPlots(strtype.map, ltrtype.map, cols=2) -->

<!-- ### Summary of Data Cleaning -->

<!-- We have applied ten separate filters to our two datasets. In the name of transparency and to highlight the bounds of our research question it is useful to review these filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes nearly 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 7,200 properties structural characteristics filters remove another 480 from consideration.  Over 1,000 of the remaining properties were never been booked and/or had outlying nightly rates and were removed.  Finally, 14 properties had either bad location data or missing information host policy. From over 20,000, we have 5,788 Airbnb properties meet the requirements of our research question. -->

<!-- For the long-term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset. -->

@@

<!-- ```{r plot_revdens, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='\\label{revdens}Comparison of Revenue Distributions'} -->

<!--   ggplot(revdens.data, aes(x=revenue, fill=tenure, color=tenure)) + -->
<!--     geom_density(alpha=.5) + -->
<!--     scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) + -->
<!--     scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) + -->
<!--     xlab('\nAnnual Revenue') + -->
<!--     scale_x_continuous(breaks=c(seq(0,60000, by=20000)), -->
<!--                        labels=c('$0', '$20k', '$40k', '$60k'))+ -->
<!--     theme(legend.position='none', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           axis.text.y = element_blank(), -->
<!--           axis.ticks.y = element_blank(), -->
<!--           axis.title.y=element_blank()) + -->
<!--     coord_cartesian(xlim=c(0, 70000)) + -->
<!--     facet_wrap(~tenure) -->

<!-- ### Comparison -->

<!-- In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing likely long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues. We first focus on using the Actual measure of net revenue for the short-term properties as discussed above.  We save discussion and analysis of the Potential revenue for the sensitivity tests later in the study.  -->

<!-- We make our imputations via hedonic price models. We model the long-term market with the long-term data and then use this model to predict or impute the likely long-term weekly rental prices for the short-term properties. We use the following model specification: -->

<!-- \begin{equation} -->
<!--   \label{eq:test} -->
<!-- ln(RentLT_{i}) = \beta_{1} BedBath_{i} + \gamma_{2} Suburb_{i} + \tau_{3} Time_{i} + \epsilon_{i} -->
<!-- \end{equation} -->

<!-- where $ln(RentLT_{i})$ is the natural log of the weekly rent, *BedBath* is the bedroom/bathroom combination, *Suburb* is a fixed effect for the location of the property, *Time* is the month of rental and $\epsilon$ a random error term. -->

<!-- Realizing that the imputed long-term rental values may be more accurate when estimated at smaller scales, we impute long-term rental rates at submarket level, specifically by type, bed/bath and geographic area. In other words, we separately estimate LTR rents for Apartments with 1bed/1bath in the city core, Apartments with 1bed/1bath in the beach submarket, etc. for all combinations of type, bed/bath and geographic area. We then combine the results from all estimates back into a single dataset that contains all of the short-term rentals.  -->

<!-- ## Results -->

<!-- We will begin our analysis by making a comparison of the short and long-term markets at the metropolitan level.  Of our sample of 5,783 short-term rental units, 1,600 or 27.7\% of those units recorded a net revenue greater than what would likely have been garnered in the long-term market.   -->

<!-- Occupancy rates are one of the most important components in determining the profitability of short-term rentals. For reference, the mean occupancy rate in the sample is 27.6\%, with a median value of 19.4\%. In \autoref{occpref} (left panel) we show the percentage of properties for which short-term rental is more profitable versus the occupancy rate over the year analyzed. We see that likelihood of short-term rental being a better option increases sharply with occupancy rate up from about 25\% up to 75\% occupancy. For occupancy rates less than 40\%, long-term rental is more likely to be preferable.  -->

<!-- As the distribution of occupancy rates is not uniform the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long-term. In the right hand panel of \autoref{occpref} (right panel) we transform the x-axis to represent the percentile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 25th percentile of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long-term one.   -->

<!-- ```{r plot_occ_metro, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{occpref}Probability of Short-Term Preference vs Occupancy Rate'} -->

<!--  ## Make plots   -->

<!--   # Raw -->
<!--   raw.occ.metro <- ggplot(rawocc.metro$data, -->
<!--                    aes(x=x.var, y=Var))+ -->
<!--     geom_point(alpha=.5, size=.4) + -->
<!--     stat_smooth(se=FALSE, size=2, span=.5) + -->
<!--     xlab('\nOccupancy Rate') + -->
<!--     scale_x_continuous(breaks=seq(0, 100, by=25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     ylab('\nSTR Preferability %') + -->
<!--     scale_y_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     theme(legend.position='none') +  -->
<!--     ggtitle('Raw Rates') +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

<!--   # Quantile -->
<!--   qtl.occ.metro <- ggplot(qtlocc.metro$data, -->
<!--                    aes(x=x.var, y=Var))+ -->
<!--     geom_point(alpha=.5, size=.4) + -->
<!--     stat_smooth(se=FALSE, size=2, span=.5) + -->
<!--     xlab('\nOccupancy Rate (Pcntl)') + -->
<!--     scale_x_continuous(breaks=seq(0, 100, by=25), -->
<!--                        labels=c('0%', '25th', '50th', '75th', '100th')) + -->
<!--     ylab('\nSTR Preferability %') + -->
<!--     scale_y_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     theme(legend.position='none') +  -->
<!--     ggtitle('Percentile Rates') +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

<!--   # Make Plot -->
<!--   ggMultiPlots(raw.occ.metro, qtl.occ.metro, cols=2) -->

<!-- Along with occupancy rates, the other important consideration is the nightly rate.  High occupancy rates with low nightly rates may not be any better than lower occupancy rate with a higher nightly rate.  In \autoref{metrohm} (left panel) we plot nightly rates versus occupancy rates.  We then divide the area of the scatter plot into 400 equal areas (20 x 20) and color each based on which is more common in the area, short-term or long-term revenue.  The transparency is related to the total count of actual market properties falling into that area. -->

<!-- The general trend is what is expected; that is that higher rates and occupancies, ceteris paribus, means greater profitability for short-term rentals.  This relationship is not a perfectly linear because the long-term rental revenue varies greatly based on the characteristics of the property.  While the left panel of \autoref{metrohm} helps show the general pattern it does leave many indeterminate areas.  In order to better delimit the situations where short-term rentals are more profitable we turn to a support vector machine (SVM) algorithm. -->

<!-- An SVM can be used as a classifier to delineate the boundaries between two (or more) classes of observations.  In this case, we delineate the boundary, \autoref{metrohm} (right panel) between situations where long-term rentals are more profitable (red) and where short-term are (green).  In order to avoid an overly broad straight line delineation we use a polynomial fitting technique. Overall, from the SVM model we see that the short-term preference begins at about 325% occupancy rate, regardless of nightly rate.  As the occupancy rate increases, the required nighly rate decreases, down to near $100 once occupancies hit 65\% or greater. -->

<!-- ```{r plot_svmr_metro, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, fig.cap='\\label{metrohm}Analysis of Raw Rates: Metro Level'} -->

<!--  ## Create the rate heat map -->

<!--   rate.hm <- abbHeatMap(str_df.ic, -->
<!--                 x.field='occ.rate', -->
<!--                 y.field='med.rate', -->
<!--                 pref.field='str.act.pref', -->
<!--                 svm=F,  -->
<!--                 alpha.count=T, -->
<!--                 add.points=T, -->
<!--                 fill.colors=c(strp.col, ltrp.col)) -->
<!--   rate.hm <- rate.hm + ylab('Nightly Rate') -->

<!--  ## Create raw SVM heatmap   -->

<!--   rate.hm.svm <- abbHeatMap(str_df.ic, -->
<!--                             x.field='occ.rate', -->
<!--                             y.field='med.rate', -->
<!--                             pref.field='str.act.pref', -->
<!--                             svm=T,  -->
<!--                             alpha.count=F, -->
<!--                             add.points=T, -->
<!--                             fill.colors=c(strp.col, ltrp.col)) -->
<!--   rate.hm.svm <- rate.hm.svm + ylab('Nightly Rate') -->

<!--   ggMultiPlots(rate.hm, rate.hm.svm, cols=2) -->

<!-- Both the occupancy rates and the nightly rates are not uniformly distributed and therefore the plots in \autoref{metrohm} may be misleading.  We convert both axis to percentiles and replot both in \autoref{metrohmp}.  Here we see a more even distibution of points across the figures.  Additionally, a more reasonable SVM breakdown occurs as well.  The right panel shows that in order for the likelihood of a property to be more profitable in short-term rental to be above 50\% the lowest occupancy rate it can have at around the 60th percentile of the market it is in.  And that only holds if the nightly rate is in the top 95th percentile of the market.  For example, if we drop the nightly rate to the median, the occupancy rate needs to be at the 75th percentile or above. -->

<!-- ```{r plot_svmq_metro, fig.width=8, fig.height=4, echo=FALSE, fig.cap='\\label{metrohmp}Analysis of Percentile Rates: Metro Level'} -->

<!--  ## Build percentile heatmap -->

<!--   qtl.hm <- abbHeatMap(str_df.ic, -->
<!--                 x.field='occ.qtl', -->
<!--                 y.field='rate.qtl', -->
<!--                 pref.field='str.act.pref', -->
<!--                 svm=F,  -->
<!--                 alpha.count=T, -->
<!--                 add.points=T, -->
<!--                 fill.colors=c(strp.col, ltrp.col)) -->
<!--   qtl.hm <- qtl.hm + ylab('Nightly Rate (Pcntl)') + xlab('Occupancy Rate (Pcntl)') -->

<!--  ## Build percentile SVM   -->

<!--   qtl.hm.svm <- abbHeatMap(str_df.ic, -->
<!--                    x.field='occ.qtl', -->
<!--                    y.field='rate.qtl', -->
<!--                    pref.field='str.act.pref', -->
<!--                    svm=T,  -->
<!--                    alpha.count=F, -->
<!--                    add.points=T, -->
<!--                    quantile=T, -->
<!--                    fill.colors=c(strp.col, ltrp.col)) -->
<!--   qtl.hm.svm <- qtl.hm.svm + ylab('Nightly Rate (Pcntl)') + xlab('Occupancy Rate (Pcntl)') -->

<!--   ## Make Plots -->

<!--   ggMultiPlots(qtl.hm, qtl.hm.svm, cols=2) -->

<!-- ### Submarkets -->

<!-- To gain a deeper understanding of the differences in short-term preference, we break the data down across a variety of characteristics and examine the preferability metrics in these submarkets. First, we bifurcate our data into houses and apartments.  Within the houses, 17.5\% of the properties are short-term preferable, while for the apartment that figure is nearly double at 30.5\%. Plotting the changes in short-term preferability against occupancy rates shows considerable similarity between houses and apartments (\autoref{smocc}).  The only measurable difference is in that for houses a property's occupancy rate needs to be 10 percentiles higher than apartments to show the same likelihood of short-term preference as an apartment. Or, in other words, a given house needs to outperform its peers by a greater extent than an apartment to achieve short-term preference.  -->

<!-- ```{r plot_occ_type, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{smocc}Preference vs Occ. Rates: By Type'} -->

<!--  ## Build raw plot -->

<!--   ptr.occ <- ggplot(rawocc.type$data, -->
<!--                     aes(x=x.var, y=Var, group=ID, color=ID)) +   -->
<!--     geom_point(alpha=.5, size=.4) + -->
<!--     stat_smooth(se=FALSE, size=2, span=.5) + -->
<!--     xlab('\nOccupancy Rate') + -->
<!--     scale_color_manual(values=c('black', 'grey60'), -->
<!--                        name='') + -->
<!--     scale_x_continuous(breaks=seq(0, 100, by=25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     ylab('\nSTR Preferability %') + -->
<!--     scale_y_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     theme(legend.position='bottom') +  -->
<!--     ggtitle('Raw Rates') +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

<!--  ## Build Percentile Plots -->

<!--   ptq.occ <- ggplot(pctocc.type$data, -->
<!--                     aes(x=x.var, y=Var, group=ID, color=ID)) +   -->
<!--     geom_point(alpha=.5, size=.4) + -->
<!--     stat_smooth(se=FALSE, size=2, span=.5) + -->
<!--     xlab('\nOccupancy Rate (Pcntl)') + -->
<!--     scale_color_manual(values=c('black', 'grey60'), -->
<!--                        name='') + -->
<!--     scale_x_continuous(breaks=seq(0, 100, by=25), -->
<!--                        labels=c('0%', '25th', '50th', '75th', '100th')) + -->
<!--     ylab('\nSTR Preferability %') + -->
<!--     scale_y_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     theme(legend.position='bottom') +  -->
<!--     ggtitle('Percentile Rates') +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

<!--   ggMultiPlots(ptr.occ, ptq.occ, cols=2) -->

<!-- Extending this analysis to include nightly rates shows a similar situation (\autoref{smhm}). For houses (left panel) there are very few combinations of occupancy rate and nightly rate that offer a greater than even likelihood of short-term preference.  As an example, a median nightly rate requires a 90th percentile occupancy rate. The market for apartment (right panel) is more forgiving and occupancy rates as low as the 60th percentile can equate to a greater than even chance of short-term prefence is coupled with nightly rates near the top of the market.  A median nightly rate requires a 70th percentile occupancy rate for short-term preference likelihood.  -->

<!-- ```{r plot_svm_type, fig.width=8, fig.height=3.5, echo=FALSE, fig.cap='\\label{smhm}Analysis of Percentile Rates: By Type'} -->

<!--  ## Make House SVM Heatmap -->

<!--   house.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type == -->
<!--                                           'House', ], -->
<!--                    x.field='occ.qtl', -->
<!--                    y.field='rate.qtl', -->
<!--                    pref.field='str.act.pref', -->
<!--                    svm=T, -->
<!--                    alpha.count=F, -->
<!--                    add.points=T, -->
<!--                    quantile=T, -->
<!--                    fill.colors=c(strp.col, ltrp.col)) + -->
<!--   theme(legend.position='none') + -->
<!--   annotate("text", x = 50, y = 50, label = "Houses", size=9) + -->
<!--   ylab('Nightly Rate (Pcntl)') + -->
<!--   xlab('Occupancy Rate (Pcntl)') -->

<!--  ## Make Apartment SVM Heat Map -->

<!--   apt.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type == -->
<!--                                           'Apartment', ], -->
<!--                    x.field='occ.qtl', -->
<!--                    y.field='rate.qtl', -->
<!--                    pref.field='str.act.pref', -->
<!--                    svm=T, -->
<!--                    alpha.count=F, -->
<!--                    add.points=T, -->
<!--                    quantile=T, -->
<!--                    fill.colors=c(strp.col, ltrp.col))+ -->
<!--   theme(legend.position='none')+ -->
<!--   annotate("text", x = 50, y = 50, label = "Apartments", size=9)+ -->
<!--   ylab('Nightly Rate (Pcntl)') + -->
<!--   xlab('Occupancy Rate (Pcntl)') -->

<!--   ggMultiPlots(house.hm.svm, apt.hm.svm, cols=2) -->

<!-- #### By Geography -->

<!-- Next, we divide the data into the five geographic sub-market shown in \autoref{pointmaps}. Within the City-Core, greater than one out of three short-term properties brought in more net revenue than they likely would have under a long-term leasing scenario.  That figure drops to one in four for beach and rural, one in six for other city location and about one in nine for suburban locations.  -->

<!-- ```{r table_geo, message=FALSE, warning=FALSE, results='asis', comment=FALSE, echo=FALSE} -->

<!--  # Make Basic comparison table -->
<!--   pref.table <- abbCreateCompTable(ic.df=str_df.ic, -->
<!--                                    split.field='geo.mrkt') -->

<!--  # Convert to wide format for output -->
<!--   pref.table.wide <- dcast(pref.table, ID ~ rev.type, value.var='Var') -->

<!--  # Limit to the extrapolated pref type -->
<!--   pref.table <- pref.table[pref.table$rev.type == 'Actual', ] -->
<!--   pref.table[ ,1] <- c('City-Core', 'City', 'Suburban', 'Rural', 'Beach') -->
<!--   pref.table[ ,2] <- paste0(sprintf('%.1f', pref.table[ ,2] * 100), '%') -->

<!--  # Fix names   -->
<!--   names(pref.table) <- c('Sub-Market', 'Short-Term Preferable') -->
<!--   options(xtable.comment = FALSE) -->

<!--  ## Make printable table -->

<!--   print(xtable(pref.table[,1:2],  -->
<!--                caption='Geographic Submarket Preferability\n', -->
<!--                digits=1, -->
<!--                align='llr'), -->
<!--         include.rownames=FALSE, -->
<!--         caption.placement='top') -->

<!-- There is a noticable variability in the relationship between occupancy rates and short-term preference across the submarkets. For raw rates (\autoref{smocc}, left panel), rural properties can acheive short-term preference with lower occupancies than the other areas. This may be due to the relatively low long-term rental rates in the more rural areas coupled with high short-term rates for homes in wine country or near the mountainous terrain east of the city. Transforming to percentiles, we see that the city-core properties can reach short-term preference at a much lower occupancy rate than the others, with suburban and city as the most difficult markets.   -->

<!-- ```{r plot_occ_geo, fig.width=9, fig.height=4, message=FALSE, warning=FALSE, echo=FALSE, fig.cap='\\label{smocc}Analysis of Percentile Rates: Geographic Submarkets'}  -->

<!--  ## Set colors -->

<!--   sm.col <- abb.col[c(1, 3, 6, 5, 2)]  -->

<!--  ## Make plot objects   -->

<!--   smr.occ <- ggplot(rawocc.geo$data, -->
<!--                     aes(x=x.var, y=Var, group=ID, color=ID)) +   -->
<!--     geom_point(alpha=.5, size=.4) + -->
<!--     stat_smooth(se=FALSE, size=1.5, span=.5) + -->
<!--     xlab('\nOccupancy Rate') + -->
<!--     scale_color_manual(values=sm.col, -->
<!--                        name='') + -->
<!--     scale_x_continuous(breaks=seq(0, 100, by=25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     ylab('\nSTR Preferability %') + -->
<!--     scale_y_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     theme(legend.position='bottom') +  -->
<!--     ggtitle('Raw Rates') +  -->
<!--     theme(plot.title = element_text(hjust = 0.5)) -->

>>>>

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)
  library(xtable)

@@

After removing property level observations that are missing price data, we find that most long-term weekly rentals are greater than \$200 and less than \$1,000 per week while most median nightly rates are greater than \$50 and less than \$500 per night. We filter the property level data accordingly. We also filter the short-term daily data to include only those observations with nightly rates between \$50 and \$500.

@@

Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs. Inner suburbs are < 10km from the CBD, Middle suburbs 10 to 20km from CBD and Outer suburbs > 20km from CBD. In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high-end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts. The specific suburb of a property is the second, or finer scale, at which the market operates.

![Suburb Categories](C:/Code/research/airbnbMelbourne/images/sub_loc_map.png)
It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

1. Rural (Outer suburbs, not Beach)
2. Suburban (Middle suburbs, not Beach)
3. City (Inner Suburbs, not Beach, not Core)
4. City-Core (Select inner suburbs with tourist activities and near CBD)
5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

We select the 18 suburbs below to represent the 'core' of the city (Table 1).  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well served by public transportation and possess abundant amenities for tourists.

```{r table_corelist, results="asis", comment=FALSE, message=FALSE, echo=FALSE}

 c.subs <- data.frame(A=core_suburbs[1:6],
                      B=core_suburbs[7:12],
                      C=core_suburbs[13:18])
 names(c.subs) <- NULL
 rownames(c.subs) <- NULL

 print(xtable(c.subs, caption = 'List of Core Suburbs\n'),
       include.rownames=F,
       caption.placement="top",
       hline.after=c(0,6))

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short- and long-term rentals, colored by submarket, are shown in .

..HOLD FOR POINT LOCATION MAPS..

(These should be done a priori and converted to .tiff or other raster format)

 To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (\autoref{smcounts}). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

.. HOLD FOR LOCATION BAR CHARTS..

### Summary of Data Cleaning

We have applied ten separate filters to our two datasets. In the name of transparency and to highlight the bounds of our research question it is useful to review these filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes nearly 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 7,200 properties structural characteristics filters remove another 480 from consideration.  Over 1,000 of the remaining properties were never been booked and/or had outlying nightly rates and were removed.  Finally, 14 properties had either bad location data or missing information host policy. From over 20,000, we have 5,788 Airbnb properties meet the requirements of our research question.

For the long-term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset.

@@

...HOLD FOR REV DENSITY COMPARISON...

In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing likely long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues. We first focus on using the Actual measure of net revenue for the short-term properties as discussed above.  We save discussion and analysis of the Potential revenue for the sensitivity tests later in the study.

We make our imputations via hedonic price models. We model the long-term market with the long-term data and then use this model to predict or impute the likely long-term weekly rental prices for the short-term properties. We use the following model specification:

\begin{equation}
  \label{eq:test}
ln(RentLT_{i}) = \beta_{1} BedBath_{i} + \gamma_{2} Suburb_{i} + \tau_{3} Time_{i} + \epsilon_{i}
\end{equation}

where $ln(RentLT_{i})$ is the natural log of the weekly rent, *BedBath* is the bedroom/bathroom combination, *Suburb* is a fixed effect for the location of the property, *Time* is the month of rental and $\epsilon$ a random error term.

Realizing that the imputed long-term rental values may be more accurate when estimated at smaller scales, we impute long-term rental rates at submarket level, specifically by type, bed/bath and geographic area. In other words, we separately estimate LTR rents for Apartments with 1bed/1bath in the city core, Apartments with 1bed/1bath in the beach submarket, etc. for all combinations of type, bed/bath and geographic area. We then combine the results from all estimates back into a single dataset that contains all of the short-term rentals.

## Results
